The Efficient Handling of Guards \\in the Design of RPython's Tracing JIT

### 2012-09-07 talk/vmil2012/paper.tex
Use package flafter for float positioning
update caption
save a line
remove todo
zitemize environment, in case we need it

\usepackage{ifthen}
\usepackage{sparklines}
\usepackage{booktabs}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{colortbl}
\usepackage{wrapfig}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{relsize}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage[pdfpagelabels=true]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{flafter}

@@

@@

%Section~\ref{sec:Evaluation} presents Figures about the absolute number of
%operations for each benchmark, and the overhead produced by the information
%stored at the different levels for the guards
In this paper we want to substantiate the aforementioned observations about guards and
describe based on them the reasoning behind their implementation in
RPython's tracing just-in-time compiler. The contributions of this paper are:
\begin{itemize}
  \item An analysis of guards in the context of RPython's tracing JIT,
  \item detailed measurements about the frequency and the
  memory overhead associated with guards, and
  \item a description about how guards are implemented in the high\-
  and low-level components of RPython's JIT and a description of the rationale behind the design.
\end{itemize}

@@

RPython can mean one of two things:
\begin{itemize}
 \item the language itself
 \item the translation toolchain used to transform RPython programs to executable units
\end{itemize}
The RPython language is a
statically typed object-oriented high-level subset of Python. The subset is chosen in such a way to make type inference possible\cite{ancona_rpython:_2007}.
The language tool-set provides
several features such as automatic memory management
and just-in-time compilation.
When writing an interpreter using RPython the
programmer only has to write the interpreter for the language she is
implementing.  The second RPython component, the translation toolchain, is used
to transform the interpreter into a C program.\footnote{
    RPython can also be used to translate programs to CLR and Java
    bytecode~\cite{ancona_rpython:_2007}, but this feature is somewhat
    experimental.
}
During the transformation process
different low level aspects suited for the target environment are automatically
added to the program such as a garbage collector and a tracing JIT compiler.
The process of inserting a tracing JIT is not fully automatic but is guided by
hints from the interpreter author.

@@

\begin{figure}
    \input{figures/example.tex}
    \caption{Example program}
    \label{fig:example}
\end{figure}

@@

\begin{figure}
    \input{figures/unopt-log.tex}
    \caption{Unoptimized trace, the line numbers in the trace correspond to the line numbers in Figure~\ref{fig:trace-log}.}
    \label{fig:unopt-trace}
\end{figure}

@@

In addition to sharing as much as possible between subsequent guards,
a compact representation of the local variables of symbolic frames is used.
Every variable in the symbolic frame is encoded using two bytes.
Two bits are used as a tag to denote where the value of the variable
comes from.
The remaining 14 bits are a payload that depends on the tag bits.

The possible sources of information are:

@@

\begin{figure}
    \input{figures/log.tex}
    \caption{Optimized trace}
    \label{fig:trace-log}
\end{figure}
% section Resume Data (end)
\section{Guards in the Backend}
\label{sec:Guards in the Backend}

\begin{figure}
\includegraphics[width=0.4\textwidth]{figures/resume_data.pdf}
\caption{The resume data for Figure~\ref{fig:trace-log}}
\label{fig:resume-data}
\end{figure}

>>>>

\usepackage{ifthen}
\usepackage{sparklines}
\usepackage{booktabs}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{colortbl}
\usepackage{wrapfig}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{relsize}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{setspace}
\usepackage[pdfpagelabels=true]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{flafter}
\usepackage{enumitem}
\setitemize{noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt}

@@

\newenvironment{zitemize}% zero - line spacing itemize environment
   {\begin{list}{--}{
   \setlength{\itemsep}{0 pt}
   \setlength{\parsep}{0 pt}
   \setlength{\topsep} {0 pt} }}% the end stuff
   {\end{list}}

@@

%Section~\ref{sec:Evaluation} presents Figures about the absolute number of
%operations for each benchmark, and the overhead produced by the information
%stored at the different levels for the guards
In this paper we want to substantiate the aforementioned observations about guards and
describe based on them the reasoning behind their implementation in
RPython's tracing just-in-time compiler. The contributions of this paper are:
\begin{itemize}
  \item An analysis of guards in the context of RPython's JIT,
  \item detailed measurements about the frequency and the
  memory overhead associated with guards, and
  \item a description about how guards are implemented in the high\-
  and low-level components of RPython's JIT and a description of the rationale behind the design.
\end{itemize}

@@

RPython can mean one of two things, the language itself and the translation
toolchain used to transform RPython programs to executable units.
The RPython language is a
statically typed object-oriented high-level subset of Python. The subset is chosen in such a way to make type inference possible\cite{ancona_rpython:_2007}.
The language tool-set provides
several features such as automatic memory management
and just-in-time compilation.
When writing an interpreter using RPython the
programmer only has to write the interpreter for the language she is
implementing.  The second RPython component, the translation toolchain, is used
to transform the interpreter into a C program.\footnote{
    RPython can also be used to translate programs to CLR and Java
    bytecode~\cite{ancona_rpython:_2007}, but this feature is somewhat
    experimental.
}
During the transformation process
different low level aspects suited for the target environment are automatically
added to the program such as a garbage collector and a tracing JIT compiler.
The process of inserting a tracing JIT is not fully automatic but is guided by
hints from the interpreter author.

@@

\begin{figure}[ht]
    \input{figures/example.tex}
    \caption{Example program}
    \label{fig:example}
\end{figure}

@@

\begin{figure}[ht]
    \input{figures/unopt-log.tex}
    \caption{Unoptimized trace, the line numbers in the trace correspond to the line numbers in Figure~\ref{fig:trace-log}.}
    \label{fig:unopt-trace}
\end{figure}

@@

In addition to sharing as much as possible between subsequent guards,
a compact representation of the local variables of symbolic frames is used.
Every variable in the symbolic frame is encoded using two bytes.
Two bits are used as a tag to denote where the value of the variable
comes from.
The remaining 14 bits are a payload that depends on the tag bits.
The possible sources of information are:

@@

\begin{figure}[ht]
    \input{figures/log.tex}
    \caption{Optimized trace}
    \label{fig:trace-log}
\end{figure}
% section Resume Data (end)
\section{Guards in the Backend}
\label{sec:Guards in the Backend}

\begin{figure}[ht]
\includegraphics[width=0.4\textwidth]{figures/resume_data.pdf}
\caption{The resume data for Figure~\ref{fig:trace-log}}
\label{fig:resume-data}
\end{figure}
